### 2017-06-23 paper/refs.tex

\begin{thebibliography}{-------}
%\providecommand{Natureexlab}[1]{#1}

\bibitem[{ESA}(1997)]{esa97}
{ESA}., Ed.
\newblock {\em {The HIPPARCOS and TYCHO catalogues. Astrometric and photometric
  star catalogues derived from the ESA HIPPARCOS Space Astrometry Mission}},
  Vol. 1200, {\em ESA Special Publication},  1997.

\bibitem[{Dehnen}(1998)]{dehnen98}
{Dehnen}, W.
\newblock {The Distribution of Nearby Stars in Velocity Space Inferred from
  HIPPARCOS Data}.
\newblock {\em AJ} {\bf 1998}, {\em 115},~2384--2396,
  \href{http://xxx.lanl.gov/abs/astro-ph/9803110}{{\normalfont
  [astro-ph/9803110]}}.

\bibitem[{Dehnen}(2000)]{dehnen00}
{Dehnen}, W.
\newblock {The Effect of the Outer Lindblad Resonance of the Galactic Bar on
  the Local Stellar Velocity Distribution}.
\newblock {\em AJ} {\bf 2000}, {\em 119},~800--812,
  \href{http://xxx.lanl.gov/abs/astro-ph/9911161}{{\normalfont
  [astro-ph/9911161]}}.

\bibitem[{York} \em{et~al.}(2000){York}, {Adelman}, {Anderson}, {Anderson},
  {Annis}, {Bahcall}, {Bakken}, {Barkhouser}, {Bastian}, {Berman}, {Boroski},
  {Bracker}, {Briegel}, {Briggs}, {Brinkmann}, {Brunner}, {Burles}, {Carey},
  {Carr}, {Castander}, {Chen}, {Colestock}, {Connolly}, {Crocker}, {Csabai},
  {Czarapata}, {Davis}, {Doi}, {Dombeck}, {Eisenstein}, {Ellman}, {Elms},
  {Evans}, {Fan}, {Federwitz}, {Fiscelli}, {Friedman}, {Frieman}, {Fukugita},
  {Gillespie}, {Gunn}, {Gurbani}, {de Haas}, {Haldeman}, {Harris}, {Hayes},
  {Heckman}, {Hennessy}, {Hindsley}, {Holm}, {Holmgren}, {Huang}, {Hull},
  {Husby}, {Ichikawa}, {Ichikawa}, {Ivezi{\'c}}, {Kent}, {Kim}, {Kinney},
  {Klaene}, {Kleinman}, {Kleinman}, {Knapp}, {Korienek}, {Kron}, {Kunszt},
  {Lamb}, {Lee}, {Leger}, {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis},
  {Loveday}, {Lucinio}, {Lupton}, {MacKinnon}, {Mannery}, {Mantsch}, {Margon},
  {McGehee}, {McKay}, {Meiksin}, {Merelli}, {Monet}, {Munn}, {Narayanan},
  {Nash}, {Neilsen}, {Neswold}, {Newberg}, {Nichol}, {Nicinski}, {Nonino},
  {Okada}, {Okamura}, {Ostriker}, {Owen}, {Pauls}, {Peoples}, {Peterson},
  {Petravick}, {Pier}, {Pope}, {Pordes}, {Prosapio}, {Rechenmacher}, {Quinn},
  {Richards}, {Richmond}, {Rivetta}, {Rockosi}, {Ruthmansdorfer}, {Sandford},
  {Schlegel}, {Schneider}, {Sekiguchi}, {Sergey}, {Shimasaku}, {Siegmund},
  {Smee}, {Smith}, {Snedden}, {Stone}, {Stoughton}, {Strauss}, {Stubbs},
  {SubbaRao}, {Szalay}, {Szapudi}, {Szokoly}, {Thakar}, {Tremonti}, {Tucker},
  {Uomoto}, {Vanden Berk}, {Vogeley}, {Waddell}, {Wang}, {Watanabe},
  {Weinberg}, {Yanny}, {Yasuda}, and {SDSS Collaboration}]{york00}
{York}, D.G.; {Adelman}, J.; {Anderson}, Jr., J.E.; {Anderson}, S.F.; {Annis},
  J.; {Bahcall}, N.A.; {Bakken}, J.A.; {Barkhouser}, R.; {Bastian}, S.;
  {Berman}, E.; {Boroski}, W.N.; {Bracker}, S.; {Briegel}, C.; {Briggs}, J.W.;
  {Brinkmann}, J.; {Brunner}, R.; {Burles}, S.; {Carey}, L.; {Carr}, M.A.;
  {Castander}, F.J.; {Chen}, B.; {Colestock}, P.L.; {Connolly}, A.J.;
  {Crocker}, J.H.; {Csabai}, I.; {Czarapata}, P.C.; {Davis}, J.E.; {Doi}, M.;
  {Dombeck}, T.; {Eisenstein}, D.; {Ellman}, N.; {Elms}, B.R.; {Evans}, M.L.;
  {Fan}, X.; {Federwitz}, G.R.; {Fiscelli}, L.; {Friedman}, S.; {Frieman},
  J.A.; {Fukugita}, M.; {Gillespie}, B.; {Gunn}, J.E.; {Gurbani}, V.K.; {de
  Haas}, E.; {Haldeman}, M.; {Harris}, F.H.; {Hayes}, J.; {Heckman}, T.M.;
  {Hennessy}, G.S.; {Hindsley}, R.B.; {Holm}, S.; {Holmgren}, D.J.; {Huang},
  C.h.; {Hull}, C.; {Husby}, D.; {Ichikawa}, S.I.; {Ichikawa}, T.;
  {Ivezi{\'c}}, {\v Z}.; {Kent}, S.; {Kim}, R.S.J.; {Kinney}, E.; {Klaene}, M.;
  {Kleinman}, A.N.; {Kleinman}, S.; {Knapp}, G.R.; {Korienek}, J.; {Kron},
  R.G.; {Kunszt}, P.Z.; {Lamb}, D.Q.; {Lee}, B.; {Leger}, R.F.; {Limmongkol},
  S.; {Lindenmeyer}, C.; {Long}, D.C.; {Loomis}, C.; {Loveday}, J.; {Lucinio},
  R.; {Lupton}, R.H.; {MacKinnon}, B.; {Mannery}, E.J.; {Mantsch}, P.M.;
  {Margon}, B.; {McGehee}, P.; {McKay}, T.A.; {Meiksin}, A.; {Merelli}, A.;
  {Monet}, D.G.; {Munn}, J.A.; {Narayanan}, V.K.; {Nash}, T.; {Neilsen}, E.;
  {Neswold}, R.; {Newberg}, H.J.; {Nichol}, R.C.; {Nicinski}, T.; {Nonino}, M.;
  {Okada}, N.; {Okamura}, S.; {Ostriker}, J.P.; {Owen}, R.; {Pauls}, A.G.;
  {Peoples}, J.; {Peterson}, R.L.; {Petravick}, D.; {Pier}, J.R.; {Pope}, A.;
  {Pordes}, R.; {Prosapio}, A.; {Rechenmacher}, R.; {Quinn}, T.R.; {Richards},
  G.T.; {Richmond}, M.W.; {Rivetta}, C.H.; {Rockosi}, C.M.; {Ruthmansdorfer},
  K.; {Sandford}, D.; {Schlegel}, D.J.; {Schneider}, D.P.; {Sekiguchi}, M.;
  {Sergey}, G.; {Shimasaku}, K.; {Siegmund}, W.A.; {Smee}, S.; {Smith}, J.A.;
  {Snedden}, S.; {Stone}, R.; {Stoughton}, C.; {Strauss}, M.A.; {Stubbs}, C.;
  {SubbaRao}, M.; {Szalay}, A.S.; {Szapudi}, I.; {Szokoly}, G.P.; {Thakar},
  A.R.; {Tremonti}, C.; {Tucker}, D.L.; {Uomoto}, A.; {Vanden Berk}, D.;
  {Vogeley}, M.S.; {Waddell}, P.; {Wang}, S.i.; {Watanabe}, M.; {Weinberg},
  D.H.; {Yanny}, B.; {Yasuda}, N.; {SDSS Collaboration}.
\newblock {The Sloan Digital Sky Survey: Technical Summary}.
\newblock {\em AJ} {\bf 2000}, {\em 120},~1579--1587,
  \href{http://xxx.lanl.gov/abs/astro-ph/0006396}{{\normalfont
  [astro-ph/0006396]}}.

\bibitem[{Stoughton} \em{et~al.}(2002){Stoughton}, {Lupton}, {Bernardi},
  {Blanton}, {Burles}, {Castander}, {Connolly}, {Eisenstein}, {Frieman},
  {Hennessy}, {Hindsley}, {Ivezi{\'c}}, {Kent}, {Kunszt}, {Lee}, {Meiksin},
  {Munn}, {Newberg}, {Nichol}, {Nicinski}, {Pier}, {Richards}, {Richmond},
  {Schlegel}, {Smith}, {Strauss}, {SubbaRao}, {Szalay}, {Thakar}, {Tucker},
  {Vanden Berk}, {Yanny}, {Adelman}, {Anderson}, {Anderson}, {Annis},
  {Bahcall}, {Bakken}, {Bartelmann}, {Bastian}, {Bauer}, {Berman},
  {B{\"o}hringer}, {Boroski}, {Bracker}, {Briegel}, {Briggs}, {Brinkmann},
  {Brunner}, {Carey}, {Carr}, {Chen}, {Christian}, {Colestock}, {Crocker},
  {Csabai}, {Czarapata}, {Dalcanton}, {Davidsen}, {Davis}, {Dehnen},
  {Dodelson}, {Doi}, {Dombeck}, {Donahue}, {Ellman}, {Elms}, {Evans}, {Eyer},
  {Fan}, {Federwitz}, {Friedman}, {Fukugita}, {Gal}, {Gillespie}, {Glazebrook},
  {Gray}, {Grebel}, {Greenawalt}, {Greene}, {Gunn}, {de Haas}, {Haiman},
  {Haldeman}, {Hall}, {Hamabe}, {Hansen}, {Harris}, {Harris}, {Harvanek},
  {Hawley}, {Hayes}, {Heckman}, {Helmi}, {Henden}, {Hogan}, {Hogg}, {Holmgren},
  {Holtzman}, {Huang}, {Hull}, {Ichikawa}, {Ichikawa}, {Johnston}, {Kauffmann},
  {Kim}, {Kimball}, {Kinney}, {Klaene}, {Kleinman}, {Klypin}, {Knapp},
  {Korienek}, {Krolik}, {Kron}, {Krzesi{\'n}ski}, {Lamb}, {Leger},
  {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis}, {Loveday}, {MacKinnon},
  {Mannery}, {Mantsch}, {Margon}, {McGehee}, {McKay}, {McLean}, {Menou},
  {Merelli}, {Mo}, {Monet}, {Nakamura}, {Narayanan}, {Nash}, {Neilsen},
  {Newman}, {Nitta}, {Odenkirchen}, {Okada}, {Okamura}, {Ostriker}, {Owen},
  {Pauls}, {Peoples}, {Peterson}, {Petravick}, {Pope}, {Pordes}, {Postman},
  {Prosapio}, {Quinn}, {Rechenmacher}, {Rivetta}, {Rix}, {Rockosi}, {Rosner},
  {Ruthmansdorfer}, {Sandford}, {Schneider}, {Scranton}, {Sekiguchi}, {Sergey},
  {Sheth}, {Shimasaku}, {Smee}, {Snedden}, {Stebbins}, {Stubbs}, {Szapudi},
  {Szkody}, {Szokoly}, {Tabachnik}, {Tsvetanov}, {Uomoto}, {Vogeley}, {Voges},
  {Waddell}, {Walterbos}, {Wang}, {Watanabe}, {Weinberg}, {White}, {White},
  {Wilhite}, {Wolfe}, {Yasuda}, {York}, {Zehavi}, and {Zheng}]{stoughton02}
{Stoughton}, C.; {Lupton}, R.H.; {Bernardi}, M.; {Blanton}, M.R.; {Burles}, S.;
  {Castander}, F.J.; {Connolly}, A.J.; {Eisenstein}, D.J.; {Frieman}, J.A.;
  {Hennessy}, G.S.; {Hindsley}, R.B.; {Ivezi{\'c}}, {\v Z}.; {Kent}, S.;
  {Kunszt}, P.Z.; {Lee}, B.C.; {Meiksin}, A.; {Munn}, J.A.; {Newberg}, H.J.;
  {Nichol}, R.C.; {Nicinski}, T.; {Pier}, J.R.; {Richards}, G.T.; {Richmond},
  M.W.; {Schlegel}, D.J.; {Smith}, J.A.; {Strauss}, M.A.; {SubbaRao}, M.;
  {Szalay}, A.S.; {Thakar}, A.R.; {Tucker}, D.L.; {Vanden Berk}, D.E.; {Yanny},
  B.; {Adelman}, J.K.; {Anderson}, Jr., J.E.; {Anderson}, S.F.; {Annis}, J.;
  {Bahcall}, N.A.; {Bakken}, J.A.; {Bartelmann}, M.; {Bastian}, S.; {Bauer},
  A.; {Berman}, E.; {B{\"o}hringer}, H.; {Boroski}, W.N.; {Bracker}, S.;
  {Briegel}, C.; {Briggs}, J.W.; {Brinkmann}, J.; {Brunner}, R.; {Carey}, L.;
  {Carr}, M.A.; {Chen}, B.; {Christian}, D.; {Colestock}, P.L.; {Crocker},
  J.H.; {Csabai}, I.; {Czarapata}, P.C.; {Dalcanton}, J.; {Davidsen}, A.F.;
  {Davis}, J.E.; {Dehnen}, W.; {Dodelson}, S.; {Doi}, M.; {Dombeck}, T.;
  {Donahue}, M.; {Ellman}, N.; {Elms}, B.R.; {Evans}, M.L.; {Eyer}, L.; {Fan},
  X.; {Federwitz}, G.R.; {Friedman}, S.; {Fukugita}, M.; {Gal}, R.;
  {Gillespie}, B.; {Glazebrook}, K.; {Gray}, J.; {Grebel}, E.K.; {Greenawalt},
  B.; {Greene}, G.; {Gunn}, J.E.; {de Haas}, E.; {Haiman}, Z.; {Haldeman}, M.;
  {Hall}, P.B.; {Hamabe}, M.; {Hansen}, B.; {Harris}, F.H.; {Harris}, H.;
  {Harvanek}, M.; {Hawley}, S.L.; {Hayes}, J.J.E.; {Heckman}, T.M.; {Helmi},
  A.; {Henden}, A.; {Hogan}, C.J.; {Hogg}, D.W.; {Holmgren}, D.J.; {Holtzman},
  J.; {Huang}, C.H.; {Hull}, C.; {Ichikawa}, S.I.; {Ichikawa}, T.; {Johnston},
  D.E.; {Kauffmann}, G.; {Kim}, R.S.J.; {Kimball}, T.; {Kinney}, E.; {Klaene},
  M.; {Kleinman}, S.J.; {Klypin}, A.; {Knapp}, G.R.; {Korienek}, J.; {Krolik},
  J.; {Kron}, R.G.; {Krzesi{\'n}ski}, J.; {Lamb}, D.Q.; {Leger}, R.F.;
  {Limmongkol}, S.; {Lindenmeyer}, C.; {Long}, D.C.; {Loomis}, C.; {Loveday},
  J.; {MacKinnon}, B.; {Mannery}, E.J.; {Mantsch}, P.M.; {Margon}, B.;
  {McGehee}, P.; {McKay}, T.A.; {McLean}, B.; {Menou}, K.; {Merelli}, A.; {Mo},
  H.J.; {Monet}, D.G.; {Nakamura}, O.; {Narayanan}, V.K.; {Nash}, T.;
  {Neilsen}, Jr., E.H.; {Newman}, P.R.; {Nitta}, A.; {Odenkirchen}, M.;
  {Okada}, N.; {Okamura}, S.; {Ostriker}, J.P.; {Owen}, R.; {Pauls}, A.G.;
  {Peoples}, J.; {Peterson}, R.S.; {Petravick}, D.; {Pope}, A.; {Pordes}, R.;
  {Postman}, M.; {Prosapio}, A.; {Quinn}, T.R.; {Rechenmacher}, R.; {Rivetta},
  C.H.; {Rix}, H.W.; {Rockosi}, C.M.; {Rosner}, R.; {Ruthmansdorfer}, K.;
  {Sandford}, D.; {Schneider}, D.P.; {Scranton}, R.; {Sekiguchi}, M.; {Sergey},
  G.; {Sheth}, R.; {Shimasaku}, K.; {Smee}, S.; {Snedden}, S.A.; {Stebbins},
  A.; {Stubbs}, C.; {Szapudi}, I.; {Szkody}, P.; {Szokoly}, G.P.; {Tabachnik},
  S.; {Tsvetanov}, Z.; {Uomoto}, A.; {Vogeley}, M.S.; {Voges}, W.; {Waddell},
  P.; {Walterbos}, R.; {Wang}, S.i.; {Watanabe}, M.; {Weinberg}, D.H.; {White},
  R.L.; {White}, S.D.M.; {Wilhite}, B.; {Wolfe}, D.; {Yasuda}, N.; {York},
  D.G.; {Zehavi}, I.; {Zheng}, W.
\newblock {Sloan Digital Sky Survey: Early Data Release}.
\newblock {\em AJ} {\bf 2002}, {\em 123},~485--548.

\bibitem[{Abazajian} \em{et~al.}(2003){Abazajian}, {Adelman-McCarthy},
  {Ag{\"u}eros}, {Allam}, {Anderson}, {Annis}, {Bahcall}, {Baldry}, {Bastian},
  {Berlind}, {Bernardi}, {Blanton}, {Blythe}, {Bochanski}, {Boroski},
  {Brewington}, {Briggs}, {Brinkmann}, {Brunner}, {Budav{\'a}ri}, {Carey},
  {Carr}, {Castander}, {Chiu}, {Collinge}, {Connolly}, {Covey}, {Csabai},
  {Dalcanton}, {Dodelson}, {Doi}, {Dong}, {Eisenstein}, {Evans}, {Fan},
  {Feldman}, {Finkbeiner}, {Friedman}, {Frieman}, {Fukugita}, {Gal},
  {Gillespie}, {Glazebrook}, {Gonzalez}, {Gray}, {Grebel}, {Grodnicki}, {Gunn},
  {Gurbani}, {Hall}, {Hao}, {Harbeck}, {Harris}, {Harris}, {Harvanek},
  {Hawley}, {Heckman}, {Helmboldt}, {Hendry}, {Hennessy}, {Hindsley}, {Hogg},
  {Holmgren}, {Holtzman}, {Homer}, {Hui}, {Ichikawa}, {Ichikawa}, {Inkmann},
  {Ivezi{\'c}}, {Jester}, {Johnston}, {Jordan}, {Jordan}, {Jorgensen},
  {Juri{\'c}}, {Kauffmann}, {Kent}, {Kleinman}, {Knapp}, {Kniazev}, {Kron},
  {Krzesi{\'n}ski}, {Kunszt}, {Kuropatkin}, {Lamb}, {Lampeitl}, {Laubscher},
  {Lee}, {Leger}, {Li}, {Lidz}, {Lin}, {Loh}, {Long}, {Loveday}, {Lupton},
  {Malik}, {Margon}, {McGehee}, {McKay}, {Meiksin}, {Miknaitis}, {Moorthy},
  {Munn}, {Murphy}, {Nakajima}, {Narayanan}, {Nash}, {Neilsen}, {Newberg},
  {Newman}, {Nichol}, {Nicinski}, {Nieto-Santisteban}, {Nitta}, {Odenkirchen},
  {Okamura}, {Ostriker}, {Owen}, {Padmanabhan}, {Peoples}, {Pier}, {Pindor},
  {Pope}, {Quinn}, {Rafikov}, {Raymond}, {Richards}, {Richmond}, {Rix},
  {Rockosi}, {Schaye}, {Schlegel}, {Schneider}, {Schroeder}, {Scranton},
  {Sekiguchi}, {Seljak}, {Sergey}, {Sesar}, {Sheldon}, {Shimasaku}, {Siegmund},
  {Silvestri}, {Sinisgalli}, {Sirko}, {Smith}, {Smol{\v c}i{\'c}}, {Snedden},
  {Stebbins}, {Steinhardt}, {Stinson}, {Stoughton}, {Strateva}, {Strauss},
  {SubbaRao}, {Szalay}, {Szapudi}, {Szkody}, {Tasca}, {Tegmark}, {Thakar},
  {Tremonti}, {Tucker}, {Uomoto}, {Vanden Berk}, {Vandenberg}, {Vogeley},
  {Voges}, {Vogt}, {Walkowicz}, {Weinberg}, {West}, {White}, {Wilhite},
  {Willman}, {Xu}, {Yanny}, {Yarger}, {Yasuda}, {Yip}, {Yocum}, {York},
  {Zakamska}, {Zehavi}, {Zheng}, {Zibetti}, and {Zucker}]{abazajian03}
{Abazajian}, K.; {Adelman-McCarthy}, J.K.; {Ag{\"u}eros}, M.A.; {Allam}, S.S.;
  {Anderson}, S.F.; {Annis}, J.; {Bahcall}, N.A.; {Baldry}, I.K.; {Bastian},
  S.; {Berlind}, A.; {Bernardi}, M.; {Blanton}, M.R.; {Blythe}, N.;
  {Bochanski}, Jr., J.J.; {Boroski}, W.N.; {Brewington}, H.; {Briggs}, J.W.;
  {Brinkmann}, J.; {Brunner}, R.J.; {Budav{\'a}ri}, T.; {Carey}, L.N.; {Carr},
  M.A.; {Castander}, F.J.; {Chiu}, K.; {Collinge}, M.J.; {Connolly}, A.J.;
  {Covey}, K.R.; {Csabai}, I.; {Dalcanton}, J.J.; {Dodelson}, S.; {Doi}, M.;
  {Dong}, F.; {Eisenstein}, D.J.; {Evans}, M.L.; {Fan}, X.; {Feldman}, P.D.;
  {Finkbeiner}, D.P.; {Friedman}, S.D.; {Frieman}, J.A.; {Fukugita}, M.; {Gal},
  R.R.; {Gillespie}, B.; {Glazebrook}, K.; {Gonzalez}, C.F.; {Gray}, J.;
  {Grebel}, E.K.; {Grodnicki}, L.; {Gunn}, J.E.; {Gurbani}, V.K.; {Hall}, P.B.;
  {Hao}, L.; {Harbeck}, D.; {Harris}, F.H.; {Harris}, H.C.; {Harvanek}, M.;
  {Hawley}, S.L.; {Heckman}, T.M.; {Helmboldt}, J.F.; {Hendry}, J.S.;
  {Hennessy}, G.S.; {Hindsley}, R.B.; {Hogg}, D.W.; {Holmgren}, D.J.;
  {Holtzman}, J.A.; {Homer}, L.; {Hui}, L.; {Ichikawa}, S.i.; {Ichikawa}, T.;
  {Inkmann}, J.P.; {Ivezi{\'c}}, {\v Z}.; {Jester}, S.; {Johnston}, D.E.;
  {Jordan}, B.; {Jordan}, W.P.; {Jorgensen}, A.M.; {Juri{\'c}}, M.;
  {Kauffmann}, G.; {Kent}, S.M.; {Kleinman}, S.J.; {Knapp}, G.R.; {Kniazev},
  A.Y.; {Kron}, R.G.; {Krzesi{\'n}ski}, J.; {Kunszt}, P.Z.; {Kuropatkin}, N.;
  {Lamb}, D.Q.; {Lampeitl}, H.; {Laubscher}, B.E.; {Lee}, B.C.; {Leger}, R.F.;
  {Li}, N.; {Lidz}, A.; {Lin}, H.; {Loh}, Y.S.; {Long}, D.C.; {Loveday}, J.;
  {Lupton}, R.H.; {Malik}, T.; {Margon}, B.; {McGehee}, P.M.; {McKay}, T.A.;
  {Meiksin}, A.; {Miknaitis}, G.A.; {Moorthy}, B.K.; {Munn}, J.A.; {Murphy},
  T.; {Nakajima}, R.; {Narayanan}, V.K.; {Nash}, T.; {Neilsen}, Jr., E.H.;
  {Newberg}, H.J.; {Newman}, P.R.; {Nichol}, R.C.; {Nicinski}, T.;
  {Nieto-Santisteban}, M.; {Nitta}, A.; {Odenkirchen}, M.; {Okamura}, S.;
  {Ostriker}, J.P.; {Owen}, R.; {Padmanabhan}, N.; {Peoples}, J.; {Pier}, J.R.;
  {Pindor}, B.; {Pope}, A.C.; {Quinn}, T.R.; {Rafikov}, R.R.; {Raymond}, S.N.;
  {Richards}, G.T.; {Richmond}, M.W.; {Rix}, H.W.; {Rockosi}, C.M.; {Schaye},
  J.; {Schlegel}, D.J.; {Schneider}, D.P.; {Schroeder}, J.; {Scranton}, R.;
  {Sekiguchi}, M.; {Seljak}, U.; {Sergey}, G.; {Sesar}, B.; {Sheldon}, E.;
  {Shimasaku}, K.; {Siegmund}, W.A.; {Silvestri}, N.M.; {Sinisgalli}, A.J.;
  {Sirko}, E.; {Smith}, J.A.; {Smol{\v c}i{\'c}}, V.; {Snedden}, S.A.;
  {Stebbins}, A.; {Steinhardt}, C.; {Stinson}, G.; {Stoughton}, C.; {Strateva},
  I.V.; {Strauss}, M.A.; {SubbaRao}, M.; {Szalay}, A.S.; {Szapudi}, I.;
  {Szkody}, P.; {Tasca}, L.; {Tegmark}, M.; {Thakar}, A.R.; {Tremonti}, C.;
  {Tucker}, D.L.; {Uomoto}, A.; {Vanden Berk}, D.E.; {Vandenberg}, J.;
  {Vogeley}, M.S.; {Voges}, W.; {Vogt}, N.P.; {Walkowicz}, L.M.; {Weinberg},
  D.H.; {West}, A.A.; {White}, S.D.M.; {Wilhite}, B.C.; {Willman}, B.; {Xu},
  Y.; {Yanny}, B.; {Yarger}, J.; {Yasuda}, N.; {Yip}, C.W.; {Yocum}, D.R.;
  {York}, D.G.; {Zakamska}, N.L.; {Zehavi}, I.; {Zheng}, W.; {Zibetti}, S.;
  {Zucker}, D.B.
\newblock {The First Data Release of the Sloan Digital Sky Survey}.
\newblock {\em AJ} {\bf 2003}, {\em 126},~2081--2086,
  \href{http://xxx.lanl.gov/abs/astro-ph/0305492}{{\normalfont
  [astro-ph/0305492]}}.

\bibitem[{Newberg} \em{et~al.}(2002){Newberg}, {Yanny}, {Rockosi}, {Grebel},
  {Rix}, {Brinkmann}, {Csabai}, {Hennessy}, {Hindsley}, {Ibata}, {Ivezi{\'c}},
  {Lamb}, {Nash}, {Odenkirchen}, {Rave}, {Schneider}, {Smith}, {Stolte}, and
  {York}]{newberg02}
{Newberg}, H.J.; {Yanny}, B.; {Rockosi}, C.; {Grebel}, E.K.; {Rix}, H.W.;
  {Brinkmann}, J.; {Csabai}, I.; {Hennessy}, G.; {Hindsley}, R.B.; {Ibata}, R.;
  {Ivezi{\'c}}, Z.; {Lamb}, D.; {Nash}, E.T.; {Odenkirchen}, M.; {Rave}, H.A.;
  {Schneider}, D.P.; {Smith}, J.A.; {Stolte}, A.; {York}, D.G.
\newblock {The Ghost of Sagittarius and Lumps in the Halo of the Milky Way}.
\newblock {\em ApJ} {\bf 2002}, {\em 569},~245--274,
  \href{http://xxx.lanl.gov/abs/astro-ph/0111095}{{\normalfont
  [astro-ph/0111095]}}.

\bibitem[{Belokurov} \em{et~al.}(2006){Belokurov}, {Zucker}, {Evans},
  {Gilmore}, {Vidrih}, {Bramich}, {Newberg}, {Wyse}, {Irwin}, {Fellhauer},
  {Hewett}, {Walton}, {Wilkinson}, {Cole}, {Yanny}, {Rockosi}, {Beers}, {Bell},
  {Brinkmann}, {Ivezi{\'c}}, and {Lupton}]{belokurov06}
{Belokurov}, V.; {Zucker}, D.B.; {Evans}, N.W.; {Gilmore}, G.; {Vidrih}, S.;
  {Bramich}, D.M.; {Newberg}, H.J.; {Wyse}, R.F.G.; {Irwin}, M.J.; {Fellhauer},
  M.; {Hewett}, P.C.; {Walton}, N.A.; {Wilkinson}, M.I.; {Cole}, N.; {Yanny},
  B.; {Rockosi}, C.M.; {Beers}, T.C.; {Bell}, E.F.; {Brinkmann}, J.;
  {Ivezi{\'c}}, {\v Z}.; {Lupton}, R.
\newblock {The Field of Streams: Sagittarius and Its Siblings}.
\newblock {\em ApJ Lett} {\bf 2006}, {\em 642},~L137--L140,
  \href{http://xxx.lanl.gov/abs/astro-ph/0605025}{{\normalfont
  [astro-ph/0605025]}}.

\bibitem[{Bullock} \em{et~al.}(2001){Bullock}, {Kravtsov}, and
  {Weinberg}]{bullock01}
{Bullock}, J.S.; {Kravtsov}, A.V.; {Weinberg}, D.H.
\newblock {Hierarchical Galaxy Formation and Substructure in the Galaxy's
  Stellar Halo}.
\newblock {\em ApJ} {\bf 2001}, {\em 548},~33--46,
  \href{http://xxx.lanl.gov/abs/astro-ph/0007295}{{\normalfont
  [astro-ph/0007295]}}.

\bibitem[{Bullock} and {Johnston}(2005)]{bullock05}
{Bullock}, J.S.; {Johnston}, K.V.
\newblock {Tracing Galaxy Formation with Stellar Halos. I. Methods}.
\newblock {\em ApJ} {\bf 2005}, {\em 635},~931--949,
  \href{http://xxx.lanl.gov/abs/astro-ph/0506467}{{\normalfont
  [astro-ph/0506467]}}.

\bibitem[{Nikolaev} \em{et~al.}(2000){Nikolaev}, {Weinberg}, {Skrutskie},
  {Cutri}, {Wheelock}, {Gizis}, and {Howard}]{nikolaev00}
{Nikolaev}, S.; {Weinberg}, M.D.; {Skrutskie}, M.F.; {Cutri}, R.M.; {Wheelock},
  S.L.; {Gizis}, J.E.; {Howard}, E.M.
\newblock {A Global Photometric Analysis of 2MASS Calibration Data}.
\newblock {\em AJ} {\bf 2000}, {\em 120},~3340--3350,
  \href{http://xxx.lanl.gov/abs/astro-ph/0008002}{{\normalfont
  [astro-ph/0008002]}}.

\bibitem[{Majewski} \em{et~al.}(2003){Majewski}, {Skrutskie}, {Weinberg}, and
  {Ostheimer}]{majewski03}
{Majewski}, S.R.; {Skrutskie}, M.F.; {Weinberg}, M.D.; {Ostheimer}, J.C.
\newblock {A Two Micron All Sky Survey View of the Sagittarius Dwarf Galaxy. I.
  Morphology of the Sagittarius Core and Tidal Arms}.
\newblock {\em ApJ} {\bf 2003}, {\em 599},~1082--1115,
  \href{http://xxx.lanl.gov/abs/astro-ph/0304198}{{\normalfont
  [astro-ph/0304198]}}.

\bibitem[{Law} and {Majewski}(2010)]{law10}
{Law}, D.R.; {Majewski}, S.R.
\newblock {The Sagittarius Dwarf Galaxy: A Model for Evolution in a Triaxial
  Milky Way Halo}.
\newblock {\em ApJ} {\bf 2010}, {\em 714},~229--254,
  \href{http://xxx.lanl.gov/abs/1003.1132}{{\normalfont [1003.1132]}}.

\bibitem[{Sheffield} \em{et~al.}(2014){Sheffield}, {Johnston}, {Majewski},
  {Damke}, {Richardson}, {Beaton}, and {Rocha-Pinto}]{sheffield14}
{Sheffield}, A.A.; {Johnston}, K.V.; {Majewski}, S.R.; {Damke}, G.;
  {Richardson}, W.; {Beaton}, R.; {Rocha-Pinto}, H.J.
\newblock {Exploring Halo Substructure with Giant Stars. XIV. The Nature of the
  Triangulum-Andromeda Stellar Features}.
\newblock {\em ApJ} {\bf 2014}, {\em 793},~62,
  \href{http://xxx.lanl.gov/abs/1407.4463}{{\normalfont [1407.4463]}}.

\bibitem[{Slater} \em{et~al.}(2014){Slater}, {Bell}, {Schlafly}, {Morganson},
  {Martin}, {Rix}, {Pe{\~n}arrubia}, {Bernard}, {Ferguson}, {Martinez-Delgado},
  {Wyse}, {Burgett}, {Chambers}, {Draper}, {Hodapp}, {Kaiser}, {Magnier},
  {Metcalfe}, {Price}, {Tonry}, {Wainscoat}, and {Waters}]{slater14}
{Slater}, C.T.; {Bell}, E.F.; {Schlafly}, E.F.; {Morganson}, E.; {Martin},
  N.F.; {Rix}, H.W.; {Pe{\~n}arrubia}, J.; {Bernard}, E.J.; {Ferguson}, A.M.N.;
  {Martinez-Delgado}, D.; {Wyse}, R.F.G.; {Burgett}, W.S.; {Chambers}, K.C.;
  {Draper}, P.W.; {Hodapp}, K.W.; {Kaiser}, N.; {Magnier}, E.A.; {Metcalfe},
  N.; {Price}, P.A.; {Tonry}, J.L.; {Wainscoat}, R.J.; {Waters}, C.
\newblock {The Complex Structure of Stars in the Outer Galactic Disk as
  Revealed by Pan-STARRS1}.
\newblock {\em ApJ} {\bf 2014}, {\em 791},~9,
  \href{http://xxx.lanl.gov/abs/1406.6368}{{\normalfont [1406.6368]}}.

\bibitem[{Martin} \em{et~al.}(2014){Martin}, {Ibata}, {Rich}, {Collins},
  {Fardal}, {Irwin}, {Lewis}, {McConnachie}, {Babul}, {Bate}, {Chapman},
  {Conn}, {Crnojevi{\'c}}, {Ferguson}, {Mackey}, {Navarro}, {Pe{\~n}arrubia},
  {Tanvir}, and {Valls-Gabaud}]{martin14}
{Martin}, N.F.; {Ibata}, R.A.; {Rich}, R.M.; {Collins}, M.L.M.; {Fardal}, M.A.;
  {Irwin}, M.J.; {Lewis}, G.F.; {McConnachie}, A.W.; {Babul}, A.; {Bate}, N.F.;
  {Chapman}, S.C.; {Conn}, A.R.; {Crnojevi{\'c}}, D.; {Ferguson}, A.M.N.;
  {Mackey}, A.D.; {Navarro}, J.F.; {Pe{\~n}arrubia}, J.; {Tanvir}, N.T.;
  {Valls-Gabaud}, D.
\newblock {The PAndAS Field of Streams: Stellar Structures in the Milky Way
  Halo toward Andromeda and Triangulum}.
\newblock {\em ApJ} {\bf 2014}, {\em 787},~19,
  \href{http://xxx.lanl.gov/abs/1403.4945}{{\normalfont [1403.4945]}}.

\bibitem[{Deason} \em{et~al.}(2014){Deason}, {Belokurov}, {Hamren}, {Koposov},
  {Gilbert}, {Beaton}, {Dorman}, {Guhathakurta}, {Majewski}, and
  {Cunningham}]{deason14}
{Deason}, A.J.; {Belokurov}, V.; {Hamren}, K.M.; {Koposov}, S.E.; {Gilbert},
  K.M.; {Beaton}, R.L.; {Dorman}, C.E.; {Guhathakurta}, P.; {Majewski}, S.R.;
  {Cunningham}, E.C.
\newblock {TriAnd and its siblings: satellites of satellites in the Milky Way
  halo}.
\newblock {\em MNRAS} {\bf 2014}, {\em 444},~3975--3985,
  \href{http://xxx.lanl.gov/abs/1407.4458}{{\normalfont [1407.4458]}}.

\bibitem[{Robin} \em{et~al.}(1992){Robin}, {Creze}, and {Mohan}]{robin92}
{Robin}, A.C.; {Creze}, M.; {Mohan}, V.
\newblock {The edge of the Galactic disk}.
\newblock {\em ApJ Lett} {\bf 1992}, {\em 400},~L25--L27,
  \href{http://xxx.lanl.gov/abs/astro-ph/9210001}{{\normalfont
  [astro-ph/9210001]}}.

\bibitem[{Morganson} \em{et~al.}(2016){Morganson}, {Conn}, {Rix}, {Bell},
  {Burgett}, {Chambers}, {Dolphin}, {Draper}, {Flewelling}, {Hodapp}, {Kaiser},
  {Magnier}, {Martin}, {Martinez-Delgado}, {Metcalfe}, {Schlafly}, {Slater},
  {Wainscoat}, and {Waters}]{Morganson:2016}
{Morganson}, E.; {Conn}, B.; {Rix}, H.W.; {Bell}, E.F.; {Burgett}, W.S.;
  {Chambers}, K.; {Dolphin}, A.; {Draper}, P.W.; {Flewelling}, H.; {Hodapp},
  K.; {Kaiser}, N.; {Magnier}, E.A.; {Martin}, N.F.; {Martinez-Delgado}, D.;
  {Metcalfe}, N.; {Schlafly}, E.F.; {Slater}, C.T.; {Wainscoat}, R.J.;
  {Waters}, C.Z.
\newblock {Mapping the Monoceros Ring in 3D with Pan-STARRS1}.
\newblock {\em ApJ} {\bf 2016}, {\em 825},~140,
  \href{http://xxx.lanl.gov/abs/1604.07501}{{\normalfont [1604.07501]}}.

\bibitem[{Yanny} \em{et~al.}(2003){Yanny}, {Newberg}, {Grebel}, {Kent},
  {Odenkirchen}, {Rockosi}, {Schlegel}, {Subbarao}, {Brinkmann}, {Fukugita},
  {Ivezic}, {Lamb}, {Schneider}, and {York}]{yanny03}
{Yanny}, B.; {Newberg}, H.J.; {Grebel}, E.K.; {Kent}, S.; {Odenkirchen}, M.;
  {Rockosi}, C.M.; {Schlegel}, D.; {Subbarao}, M.; {Brinkmann}, J.; {Fukugita},
  M.; {Ivezic}, {\v Z}.; {Lamb}, D.Q.; {Schneider}, D.P.; {York}, D.G.
\newblock {A Low-Latitude Halo Stream around the Milky Way}.
\newblock {\em ApJ} {\bf 2003}, {\em 588},~824--841,
  \href{http://xxx.lanl.gov/abs/astro-ph/0301029}{{\normalfont
  [astro-ph/0301029]}}.

\bibitem[{Ibata} \em{et~al.}(2003){Ibata}, {Irwin}, {Lewis}, {Ferguson}, and
  {Tanvir}]{ibata03}
{Ibata}, R.A.; {Irwin}, M.J.; {Lewis}, G.F.; {Ferguson}, A.M.N.; {Tanvir}, N.
\newblock {One ring to encompass them all: a giant stellar structure that
  surrounds the Galaxy}.
\newblock {\em MNRAS} {\bf 2003}, {\em 340},~L21--L27,
  \href{http://xxx.lanl.gov/abs/astro-ph/0301067}{{\normalfont
  [astro-ph/0301067]}}.

\bibitem[{Rocha-Pinto} \em{et~al.}(2003){Rocha-Pinto}, {Majewski}, {Skrutskie},
  and {Crane}]{rochapinto03}
{Rocha-Pinto}, H.J.; {Majewski}, S.R.; {Skrutskie}, M.F.; {Crane}, J.D.
\newblock {Tracing the Galactic Anticenter Stellar Stream with 2MASS M Giants}.
\newblock {\em ApJ Lett} {\bf 2003}, {\em 594},~L115--L118.

\bibitem[{Rocha-Pinto} \em{et~al.}(2004){Rocha-Pinto}, {Majewski}, {Skrutskie},
  {Crane}, and {Patterson}]{rochapinto04}
{Rocha-Pinto}, H.J.; {Majewski}, S.R.; {Skrutskie}, M.F.; {Crane}, J.D.;
  {Patterson}, R.J.
\newblock {Exploring Halo Substructure with Giant Stars: A Diffuse Star Cloud
  or Tidal Debris around the Milky Way in Triangulum-Andromeda}.
\newblock {\em ApJ} {\bf 2004}, {\em 615},~732--737,
  \href{http://xxx.lanl.gov/abs/astro-ph/0405437}{{\normalfont
  [astro-ph/0405437]}}.

\bibitem[{Majewski} \em{et~al.}(2004){Majewski}, {Ostheimer}, {Rocha-Pinto},
  {Patterson}, {Guhathakurta}, and {Reitzel}]{majewski04}
{Majewski}, S.R.; {Ostheimer}, J.C.; {Rocha-Pinto}, H.J.; {Patterson}, R.J.;
  {Guhathakurta}, P.; {Reitzel}, D.
\newblock {Detection of the Main-Sequence Turnoff of a Newly Discovered Milky
  Way Halo Structure in the Triangulum-Andromeda Region}.
\newblock {\em ApJ} {\bf 2004}, {\em 615},~738--743,
  \href{http://xxx.lanl.gov/abs/astro-ph/0406221}{{\normalfont
  [astro-ph/0406221]}}.

\bibitem[{Martin} \em{et~al.}(2007){Martin}, {Ibata}, and {Irwin}]{martin07}
{Martin}, N.F.; {Ibata}, R.A.; {Irwin}, M.
\newblock {Galactic Halo Stellar Structures in the Triangulum-Andromeda
  Region}.
\newblock {\em ApJ Lett} {\bf 2007}, {\em 668},~L123--L126,
  \href{http://xxx.lanl.gov/abs/astro-ph/0703506}{{\normalfont
  [astro-ph/0703506]}}.

\bibitem[{Sharma} and {Johnston}(2009)]{sharma09}
{Sharma}, S.; {Johnston}, K.V.
\newblock {A Group Finding Algorithm for Multidimensional Data Sets}.
\newblock {\em ApJ} {\bf 2009}, {\em 703},~1061--1077.

\bibitem[{Sharma} \em{et~al.}(2010){Sharma}, {Johnston}, {Majewski},
  {Mu{\~n}oz}, {Carlberg}, and {Bullock}]{sharma10}
{Sharma}, S.; {Johnston}, K.V.; {Majewski}, S.R.; {Mu{\~n}oz}, R.R.;
  {Carlberg}, J.K.; {Bullock}, J.
\newblock {Group Finding in the Stellar Halo Using M-giants in the Two Micron
  All Sky Survey: An Extended View of the Pisces Overdensity?}
\newblock {\em ApJ} {\bf 2010}, {\em 722},~750--759,
  \href{http://xxx.lanl.gov/abs/1009.0924}{{\normalfont [1009.0924]}}.

\bibitem[{Pe{\~n}arrubia} \em{et~al.}(2005){Pe{\~n}arrubia},
  {Mart{\'{\i}}nez-Delgado}, {Rix}, {G{\'o}mez-Flechoso}, {Munn}, {Newberg},
  {Bell}, {Yanny}, {Zucker}, and {Grebel}]{penarrubia05}
{Pe{\~n}arrubia}, J.; {Mart{\'{\i}}nez-Delgado}, D.; {Rix}, H.W.;
  {G{\'o}mez-Flechoso}, M.A.; {Munn}, J.; {Newberg}, H.; {Bell}, E.F.; {Yanny},
  B.; {Zucker}, D.; {Grebel}, E.K.
\newblock {A Comprehensive Model for the Monoceros Tidal Stream}.
\newblock {\em ApJ} {\bf 2005}, {\em 626},~128--144,
  \href{http://xxx.lanl.gov/abs/astro-ph/0410448}{{\normalfont
  [astro-ph/0410448]}}.

\bibitem[{Momany} \em{et~al.}(2004){Momany}, {Zaggia}, {Bonifacio}, {Piotto},
  {De Angeli}, {Bedin}, and {Carraro}]{momany04}
{Momany}, Y.; {Zaggia}, S.R.; {Bonifacio}, P.; {Piotto}, G.; {De Angeli}, F.;
  {Bedin}, L.R.; {Carraro}, G.
\newblock {Probing the Canis Major stellar over-density as due to the Galactic
  warp}.
\newblock {\em A\&A} {\bf 2004}, {\em 421},~L29--L32,
  \href{http://xxx.lanl.gov/abs/astro-ph/0405526}{{\normalfont
  [astro-ph/0405526]}}.

\bibitem[{Momany} \em{et~al.}(2006){Momany}, {Zaggia}, {Gilmore}, {Piotto},
  {Carraro}, {Bedin}, and {de Angeli}]{momany06}
{Momany}, Y.; {Zaggia}, S.; {Gilmore}, G.; {Piotto}, G.; {Carraro}, G.;
  {Bedin}, L.R.; {de Angeli}, F.
\newblock {Outer structure of the Galactic warp and flare: explaining the Canis
  Major over-density}.
\newblock {\em A\&A} {\bf 2006}, {\em 451},~515--538,
  \href{http://xxx.lanl.gov/abs/astro-ph/0603385}{{\normalfont
  [astro-ph/0603385]}}.

\bibitem[{Kazantzidis} \em{et~al.}(2008){Kazantzidis}, {Bullock}, {Zentner},
  {Kravtsov}, and {Moustakas}]{kazantzidis08}
{Kazantzidis}, S.; {Bullock}, J.S.; {Zentner}, A.R.; {Kravtsov}, A.V.;
  {Moustakas}, L.A.
\newblock {Cold Dark Matter Substructure and Galactic Disks. I. Morphological
  Signatures of Hierarchical Satellite Accretion}.
\newblock {\em ApJ} {\bf 2008}, {\em 688},~254--276,
  \href{http://xxx.lanl.gov/abs/0708.1949}{{\normalfont [0708.1949]}}.

\bibitem[{Younger} \em{et~al.}(2008){Younger}, {Besla}, {Cox}, {Hernquist},
  {Robertson}, and {Willman}]{younger08}
{Younger}, J.D.; {Besla}, G.; {Cox}, T.J.; {Hernquist}, L.; {Robertson}, B.;
  {Willman}, B.
\newblock {On the Origin of Dynamically Cold Rings around the Milky Way}.
\newblock {\em ApJ Lett} {\bf 2008}, {\em 676},~L21,
  \href{http://xxx.lanl.gov/abs/0802.0872}{{\normalfont [0802.0872]}}.

\bibitem[{Purcell} \em{et~al.}(2011){Purcell}, {Bullock}, {Tollerud}, {Rocha},
  and {Chakrabarti}]{purcell11}
{Purcell}, C.W.; {Bullock}, J.S.; {Tollerud}, E.J.; {Rocha}, M.; {Chakrabarti},
  S.
\newblock {The Sagittarius impact as an architect of spirality and outer rings
  in the Milky Way}.
\newblock {\em Nature} {\bf 2011}, {\em 477},~301--303,
  \href{http://xxx.lanl.gov/abs/1109.2918}{{\normalfont
  [arXiv:astro-ph.GA/1109.2918]}}.

\bibitem[{Xu} \em{et~al.}(2015){Xu}, {Newberg}, {Carlin}, {Liu}, {Deng}, {Li},
  {Sch{\"o}nrich}, and {Yanny}]{xu15}
{Xu}, Y.; {Newberg}, H.J.; {Carlin}, J.L.; {Liu}, C.; {Deng}, L.; {Li}, J.;
  {Sch{\"o}nrich}, R.; {Yanny}, B.
\newblock {Rings and Radial Waves in the Disk of the Milky Way}.
\newblock {\em ApJ} {\bf 2015}, {\em 801},~105,
  \href{http://xxx.lanl.gov/abs/1503.00257}{{\normalfont [1503.00257]}}.

\bibitem[{G{\'o}mez} \em{et~al.}(2016){G{\'o}mez}, {White}, {Marinacci},
  {Slater}, {Grand}, {Springel}, and {Pakmor}]{gomez16}
{G{\'o}mez}, F.A.; {White}, S.D.M.; {Marinacci}, F.; {Slater}, C.T.; {Grand},
  R.J.J.; {Springel}, V.; {Pakmor}, R.
\newblock {A fully cosmological model of a Monoceros-like ring}.
\newblock {\em MNRAS} {\bf 2016}, {\em 456},~2779--2793,
  \href{http://xxx.lanl.gov/abs/1509.08459}{{\normalfont [1509.08459]}}.

\bibitem[{Price-Whelan} \em{et~al.}(2015){Price-Whelan}, {Johnston},
  {Sheffield}, {Laporte}, and {Sesar}]{pricewhelan15}
{Price-Whelan}, A.M.; {Johnston}, K.V.; {Sheffield}, A.A.; {Laporte}, C.F.P.;
  {Sesar}, B.
\newblock {A reinterpretation of the Triangulum-Andromeda stellar clouds: a
  population of halo stars kicked out of the Galactic disc}.
\newblock {\em MNRAS} {\bf 2015}, {\em 452},~676--685,
  \href{http://xxx.lanl.gov/abs/1503.08780}{{\normalfont [1503.08780]}}.

\bibitem[{Koposov} \em{et~al.}(2010){Koposov}, {Rix}, and {Hogg}]{koposov10}
{Koposov}, S.E.; {Rix}, H.W.; {Hogg}, D.W.
\newblock {Constraining the Milky Way Potential with a Six-Dimensional
  Phase-Space Map of the GD-1 Stellar Stream}.
\newblock {\em ApJ} {\bf 2010}, {\em 712},~260--273,
  \href{http://xxx.lanl.gov/abs/0907.1085}{{\normalfont
  [arXiv:astro-ph.GA/0907.1085]}}.

\bibitem[{K{\"u}pper} \em{et~al.}(2015){K{\"u}pper}, {Balbinot}, {Bonaca},
  {Johnston}, {Hogg}, {Kroupa}, and {Santiago}]{kuepper15}
{K{\"u}pper}, A.H.W.; {Balbinot}, E.; {Bonaca}, A.; {Johnston}, K.V.; {Hogg},
  D.W.; {Kroupa}, P.; {Santiago}, B.X.
\newblock {Globular Cluster Streams as Galactic High-Precision
  Scales --- the Poster Child Palomar 5}.
\newblock {\em ApJ} {\bf 2015}, {\em 803},~80,
  \href{http://xxx.lanl.gov/abs/1502.02658}{{\normalfont [1502.02658]}}.

\bibitem[{Bovy} \em{et~al.}(2016){Bovy}, {Bahmanyar}, {Fritz}, and
  {Kallivayalil}]{bovy16}
{Bovy}, J.; {Bahmanyar}, A.; {Fritz}, T.K.; {Kallivayalil}, N.
\newblock {The Shape of the Inner Milky Way Halo from Observations of the Pal 5
  and GD--1 Stellar Streams}.
\newblock {\em ApJ} {\bf 2016}, {\em 833},~31,
  \href{http://xxx.lanl.gov/abs/1609.01298}{{\normalfont [1609.01298]}}.

\bibitem[{Johnston} \em{et~al.}(2008){Johnston}, {Bullock}, {Sharma}, {Font},
  {Robertson}, and {Leitner}]{johnston08}
{Johnston}, K.V.; {Bullock}, J.S.; {Sharma}, S.; {Font}, A.; {Robertson}, B.E.;
  {Leitner}, S.N.
\newblock {Tracing Galaxy Formation with Stellar Halos. II. Relating
  Substructure in Phase and Abundance Space to Accretion Histories}.
\newblock {\em ApJ} {\bf 2008}, {\em 689},~936--957,
  \href{http://xxx.lanl.gov/abs/0807.3911}{{\normalfont [0807.3911]}}.

\bibitem[{Drake} \em{et~al.}(2014){Drake}, {Graham}, {Djorgovski}, {Catelan},
  {Mahabal}, {Torrealba}, {Garc{\'{\i}}a-{\'A}lvarez}, {Donalek}, {Prieto},
  {Williams}, {Larson}, {Christen sen}, {Belokurov}, {Koposov}, {Beshore},
  {Boattini}, {Gibbs}, {Hill}, {Kowalski}, {Johnson}, and {Shelly}]{drake14}
{Drake}, A.J.; {Graham}, M.J.; {Djorgovski}, S.G.; {Catelan}, M.; {Mahabal},
  A.A.; {Torrealba}, G.; {Garc{\'{\i}}a-{\'A}lvarez}, D.; {Donalek}, C.;
  {Prieto}, J.L.; {Williams}, R.; {Larson}, S.; {Christen sen}, E.;
  {Belokurov}, V.; {Koposov}, S.E.; {Beshore}, E.; {Boattini}, A.; {Gibbs}, A.;
  {Hill}, R.; {Kowalski}, R.; {Johnson}, J.; {Shelly}, F.
\newblock {The Catalina Surveys Periodic Variable Star Catalog}.
\newblock {\em ApJ Supp} {\bf 2014}, {\em 213},~9,
  \href{http://xxx.lanl.gov/abs/1405.4290}{{\normalfont
  [arXiv:astro-ph.SR/1405.4290]}}.

\bibitem[{Kirby} \em{et~al.}(2011){Kirby}, {Lanfranchi}, {Simon}, {Cohen}, and
  {Guhathakurta}]{kirby11}
{Kirby}, E.N.; {Lanfranchi}, G.A.; {Simon}, J.D.; {Cohen}, J.G.;
  {Guhathakurta}, P.
\newblock {Multi-element Abundance Measurements from Medium-resolution Spectra.
  III. Metallicity Distributions of Milky Way Dwarf Satellite Galaxies}.
\newblock {\em ApJ} {\bf 2011}, {\em 727},~78,
  \href{http://xxx.lanl.gov/abs/1011.4937}{{\normalfont
  [arXiv:astro-ph.GA/1011.4937]}}.

\bibitem[{Amrose} and {Mckay}(2001)]{amrose01}
{Amrose}, S.; {Mckay}, T.
\newblock {A Calculation of the Mean Local RR Lyrae Space Density Using ROTSE}.
\newblock {\em ApJ Lett} {\bf 2001}, {\em 560},~L151--L154.

\bibitem[{Chou} \em{et~al.}(2010{Natureexlab{a}}){Chou}, {Majewski}, {Cunha},
  {Smith}, {Patterson}, and {Mart{\'{\i}}nez-Delgado}]{chou2010b}
{Chou}, M.Y.; {Majewski}, S.R.; {Cunha}, K.; {Smith}, V.V.; {Patterson}, R.J.;
  {Mart{\'{\i}}nez-Delgado}, D.
\newblock {The Chemical Evolution of the Monoceros Ring/Galactic Anticenter
  Stellar Structure}.
\newblock {\em ApJ Lett} {\bf 2010}, {\em 720},~L5--L10,
  \href{http://xxx.lanl.gov/abs/1007.1056}{{\normalfont [1007.1056]}}.

\bibitem[{Chou} \em{et~al.}(2010{Natureexlab{b}}){Chou}, {Cunha}, {Majewski},
  {Smith}, {Patterson}, {Mart{\'{\i}}nez-Delgado}, and {Geisler}]{chou2010a}
{Chou}, M.Y.; {Cunha}, K.; {Majewski}, S.R.; {Smith}, V.V.; {Patterson}, R.J.;
  {Mart{\'{\i}}nez-Delgado}, D.; {Geisler}, D.
\newblock {A Two Micron All Sky Survey View of the Sagittarius Dwarf Galaxy.
  VI. s-Process and Titanium Abundance Variations Along the Sagittarius
  Stream}.
\newblock {\em ApJ} {\bf 2010}, {\em 708},~1290--1309,
  \href{http://xxx.lanl.gov/abs/0911.4364}{{\normalfont [0911.4364]}}.

\bibitem[{Sbordone} \em{et~al.}(2005){Sbordone}, {Bonifacio}, {Marconi},
  {Zaggia}, and {Buonanno}]{sbordone2005}
{Sbordone}, L.; {Bonifacio}, P.; {Marconi}, G.; {Zaggia}, S.; {Buonanno}, R.
\newblock {UVES observations of the Canis Major overdensity}.
\newblock {\em A\&A} {\bf 2005}, {\em 430},~L13--L16,
  \href{http://xxx.lanl.gov/abs/astro-ph/0412098}{{\normalfont
  [astro-ph/0412098]}}.

\bibitem[Vogt et al.(1994)]{vogt1994} Vogt, S.~S., and 26 colleagues 1994.\ HIRES: the high-resolution echelle spectrometer on the Keck 10-m Telescope.\ Instrumentation in Astronomy VIII 2198, 362.

\bibitem[{Bergemann}(2011)]{bergemann2011}
{Bergemann}, M.
\newblock {Ionization balance of Ti in the photospheres of the Sun and four
  late-type stars}.
\newblock {\em MNRAS} {\bf 2011}, {\em 413},~2184--2198,
  \href{http://xxx.lanl.gov/abs/1101.0828}{{\normalfont
  [arXiv:astro-ph.SR/1101.0828]}}.

\bibitem[{Bergemann} \em{et~al.}(2012){Bergemann}, {Lind}, {Collet}, {Magic},
  and {Asplund}]{bergemann2012}
{Bergemann}, M.; {Lind}, K.; {Collet}, R.; {Magic}, Z.; {Asplund}, M.
\newblock {Non-LTE line formation of Fe in late-type stars - I. Standard stars
  with 1D and 3D model atmospheres}.
\newblock {\em MNRAS} {\bf 2012}, {\em 427},~27--49,
  \href{http://xxx.lanl.gov/abs/1207.2455}{{\normalfont
  [arXiv:astro-ph.SR/1207.2455]}}.

\bibitem[{Widrow} \em{et~al.}(2012){Widrow}, {Gardner}, {Yanny}, {Dodelson},
  and {Chen}]{widrow12}
{Widrow}, L.M.; {Gardner}, S.; {Yanny}, B.; {Dodelson}, S.; {Chen}, H.Y.
\newblock {Galactoseismology: Discovery of Vertical Waves in the Galactic
  Disk}.
\newblock {\em ApJ Lett} {\bf 2012}, {\em 750},~L41,
  \href{http://xxx.lanl.gov/abs/1203.6861}{{\normalfont [1203.6861]}}.

\bibitem[{Yanny} and {Gardner}(2013)]{yanny13}
{Yanny}, B.; {Gardner}, S.
\newblock {The Stellar Number Density Distribution in the Local Solar
  Neighborhood is North-South Asymmetric}.
\newblock {\em ApJ} {\bf 2013}, {\em 777},~91,
  \href{http://xxx.lanl.gov/abs/1309.2300}{{\normalfont [1309.2300]}}.

\bibitem[{Williams} \em{et~al.}(2013){Williams}, {Steinmetz}, {Binney},
  {Siebert}, {Enke}, {Famaey}, {Minchev}, {de Jong}, {Boeche}, {Freeman},
  {Bienaym{\'e}}, {Bland-Hawthorn}, {Gibson}, {Gilmore}, {Grebel}, {Helmi},
  {Kordopatis}, {Munari}, {Navarro}, {Parker}, {Reid}, {Seabroke}, {Sharma},
  {Siviero}, {Watson}, {Wyse}, and {Zwitter}]{williams13}
{Williams}, M.E.K.; {Steinmetz}, M.; {Binney}, J.; {Siebert}, A.; {Enke}, H.;
  {Famaey}, B.; {Minchev}, I.; {de Jong}, R.S.; {Boeche}, C.; {Freeman}, K.C.;
  {Bienaym{\'e}}, O.; {Bland-Hawthorn}, J.; {Gibson}, B.K.; {Gilmore}, G.F.;
  {Grebel}, E.K.; {Helmi}, A.; {Kordopatis}, G.; {Munari}, U.; {Navarro}, J.F.;
  {Parker}, Q.A.; {Reid}, W.; {Seabroke}, G.M.; {Sharma}, S.; {Siviero}, A.;
  {Watson}, F.G.; {Wyse}, R.F.G.; {Zwitter}, T.
\newblock {The wobbly Galaxy: kinematics north and south with RAVE red-clump
  giants}.
\newblock {\em MNRAS} {\bf 2013}, {\em 436},~101--121,
  \href{http://xxx.lanl.gov/abs/1302.2468}{{\normalfont [1302.2468]}}.

\bibitem[{Cui} \em{et~al.}(2012){Cui}, {Zhao}, {Chu}, {Li}, {Li}, {Zhang},
  {Su}, {Yao}, {Wang}, {Xing}, {Li}, {Zhu}, {Wang}, {Gu}, {Luo}, {Xu}, {Zhang},
  {Liu}, {Zhang}, {Yang}, {Cao}, {Chen}, {Chen}, {Chen}, {Chen}, {Chu}, {Feng},
  {Gong}, {Hou}, {Hu}, {Hu}, {Hu}, {Jia}, {Jiang}, {Jiang}, {Jiang}, {Jin},
  {Li}, {Li}, {Li}, {Liu}, {Liu}, {Lu}, {Mao}, {Men}, {Qi}, {Qi}, {Shi},
  {Tang}, {Tao}, {Wang}, {Wang}, {Wang}, {Wang}, {Wang}, {Wang}, {Wang},
  {Wang}, {Wang}, {Wang}, {Wang}, {Wang}, {Xu}, {Xu}, {Yang}, {Yu}, {Yuan},
  {Yuan}, {Zhai}, {Zhang}, {Zhang}, {Zhang}, {Zhao}, {Zhou}, {Zhou}, {Zhu}, and
  {Zou}]{cui12}
{Cui}, X.Q.; {Zhao}, Y.H.; {Chu}, Y.Q.; {Li}, G.P.; {Li}, Q.; {Zhang}, L.P.;
  {Su}, H.J.; {Yao}, Z.Q.; {Wang}, Y.N.; {Xing}, X.Z.; {Li}, X.N.; {Zhu}, Y.T.;
  {Wang}, G.; {Gu}, B.Z.; {Luo}, A.L.; {Xu}, X.Q.; {Zhang}, Z.C.; {Liu}, G.R.;
  {Zhang}, H.T.; {Yang}, D.H.; {Cao}, S.Y.; {Chen}, H.Y.; {Chen}, J.J.; {Chen},
  K.X.; {Chen}, Y.; {Chu}, J.R.; {Feng}, L.; {Gong}, X.F.; {Hou}, Y.H.; {Hu},
  H.Z.; {Hu}, N.S.; {Hu}, Z.W.; {Jia}, L.; {Jiang}, F.H.; {Jiang}, X.; {Jiang},
  Z.B.; {Jin}, G.; {Li}, A.H.; {Li}, Y.; {Li}, Y.P.; {Liu}, G.Q.; {Liu}, Z.G.;
  {Lu}, W.Z.; {Mao}, Y.D.; {Men}, L.; {Qi}, Y.J.; {Qi}, Z.X.; {Shi}, H.M.;
  {Tang}, Z.H.; {Tao}, Q.S.; {Wang}, D.Q.; {Wang}, D.; {Wang}, G.M.; {Wang},
  H.; {Wang}, J.N.; {Wang}, J.; {Wang}, J.L.; {Wang}, J.P.; {Wang}, L.; {Wang},
  S.Q.; {Wang}, Y.; {Wang}, Y.F.; {Xu}, L.Z.; {Xu}, Y.; {Yang}, S.H.; {Yu}, Y.;
  {Yuan}, H.; {Yuan}, X.Y.; {Zhai}, C.; {Zhang}, J.; {Zhang}, Y.X.; {Zhang},
  Y.; {Zhao}, M.; {Zhou}, F.; {Zhou}, G.H.; {Zhu}, J.; {Zou}, S.C.
\newblock {The Large Sky Area Multi-Object Fiber Spectroscopic Telescope
  (LAMOST)}.
\newblock {\em Research in Astronomy and Astrophysics} {\bf 2012}, {\em
  12},~1197--1242.

\bibitem[{Zhao} \em{et~al.}(2012){Zhao}, {Zhao}, {Chu}, {Jing}, and
  {Deng}]{zhao12}
{Zhao}, G.; {Zhao}, Y.H.; {Chu}, Y.Q.; {Jing}, Y.P.; {Deng}, L.C.
\newblock {LAMOST spectral survey {---} An overview}.
\newblock {\em Research in Astronomy and Astrophysics} {\bf 2012}, {\em
  12},~723--734.

\bibitem[{Carlin} \em{et~al.}(2013){Carlin}, {DeLaunay}, {Newberg}, {Deng},
  {Gole}, {Grabowski}, {Jin}, {Liu}, {Liu}, {Luo}, {Yuan}, {Zhang}, {Zhao}, and
  {Zhao}]{carlin13}
{Carlin}, J.L.; {DeLaunay}, J.; {Newberg}, H.J.; {Deng}, L.; {Gole}, D.;
  {Grabowski}, K.; {Jin}, G.; {Liu}, C.; {Liu}, X.; {Luo}, A.L.; {Yuan}, H.;
  {Zhang}, H.; {Zhao}, G.; {Zhao}, Y.
\newblock {Substructure in Bulk Velocities of Milky Way Disk Stars}.
\newblock {\em ApJ Lett} {\bf 2013}, {\em 777},~L5,
  \href{http://xxx.lanl.gov/abs/1309.6314}{{\normalfont [1309.6314]}}.

\bibitem[{Widrow} \em{et~al.}(2014){Widrow}, {Barber}, {Chequers}, and
  {Cheng}]{widrow14}
{Widrow}, L.M.; {Barber}, J.; {Chequers}, M.H.; {Cheng}, E.
\newblock {Bending and breathing modes of the Galactic disc}.
\newblock {\em MNRAS} {\bf 2014}, {\em 440},~1971--1981,
  \href{http://xxx.lanl.gov/abs/1404.4069}{{\normalfont [1404.4069]}}.

\bibitem[{Kaiser} \em{et~al.}(2010){Kaiser}, {Burgett}, {Chambers}, {Denneau},
  {Heasley}, {Jedicke}, {Magnier}, {Morgan}, {Onaka}, and {Tonry}]{kaiser10}
{Kaiser}, N.; {Burgett}, W.; {Chambers}, K.; {Denneau}, L.; {Heasley}, J.;
  {Jedicke}, R.; {Magnier}, E.; {Morgan}, J.; {Onaka}, P.; {Tonry}, J.
\newblock {The Pan-STARRS wide-field optical/NIR imaging survey}.
\newblock  Ground-based and Airborne Telescopes III,  2010, Vol. 7733, p. 77330E.

\bibitem[{Weinberg} and {Blitz}(2006)]{weinberg06}
{Weinberg}, M.D.; {Blitz}, L.
\newblock {A Magellanic Origin for the Warp of the Galaxy}.
\newblock {\em ApJ Lett} {\bf 2006}, {\em 641},~L33--L36,
  \href{http://xxx.lanl.gov/abs/astro-ph/0601694}{{\normalfont
  [astro-ph/0601694]}}.

\bibitem[{G{\'o}mez} \em{et~al.}(2013){G{\'o}mez}, {Minchev}, {O'Shea},
  {Beers}, {Bullock}, and {Purcell}]{gomez13}
{G{\'o}mez}, F.A.; {Minchev}, I.; {O'Shea}, B.W.; {Beers}, T.C.; {Bullock},
  J.S.; {Purcell}, C.W.
\newblock {Vertical density waves in the Milky Way disc induced by the
  Sagittarius dwarf galaxy}.
\newblock {\em MNRAS} {\bf 2013}, {\em 429},~159--164,
  \href{http://xxx.lanl.gov/abs/1207.3083}{{\normalfont [1207.3083]}}.

\bibitem[{Carlin} \em{et~al.}(2010){Carlin}, {Casetti-Dinescu}, {Grillmair},
  {Majewski}, and {Girard}]{carlin10}
{Carlin}, J.L.; {Casetti-Dinescu}, D.I.; {Grillmair}, C.J.; {Majewski}, S.R.;
  {Girard}, T.M.
\newblock {Kinematics in Kapteyn's Selected Area 76: Orbital Motions Within the
  Highly Substructured Anticenter Stream}.
\newblock {\em ApJ} {\bf 2010}, {\em 725},~2290--2311,
  \href{http://xxx.lanl.gov/abs/1010.5257}{{\normalfont [1010.5257]}}.

\bibitem[{Li} \em{et~al.}(2012){Li}, {Newberg}, {Carlin}, {Deng}, {Newby},
  {Willett}, {Xu}, and {Luo}]{li12}
{Li}, J.; {Newberg}, H.J.; {Carlin}, J.L.; {Deng}, L.; {Newby}, M.; {Willett},
  B.A.; {Xu}, Y.; {Luo}, Z.
\newblock {On Rings and Streams in the Galactic Anti-Center}.
\newblock {\em ApJ} {\bf 2012}, {\em 757},~151,
  \href{http://xxx.lanl.gov/abs/1206.3842}{{\normalfont [1206.3842]}}.

\bibitem[{Ferguson} \em{et~al.}(2002){Ferguson}, {Irwin}, {Ibata}, {Lewis}, and
  {Tanvir}]{ferguson02}
{Ferguson}, A.M.N.; {Irwin}, M.J.; {Ibata}, R.A.; {Lewis}, G.F.; {Tanvir}, N.R.
\newblock {Evidence for Stellar Substructure in the Halo and Outer Disk of
  M31}.
\newblock {\em AJ} {\bf 2002}, {\em 124},~1452--1463,
  \href{http://xxx.lanl.gov/abs/astro-ph/0205530}{{\normalfont
  [astro-ph/0205530]}}.

\bibitem[{Ibata} \em{et~al.}(2005){Ibata}, {Chapman}, {Ferguson}, {Lewis},
  {Irwin}, and {Tanvir}]{ibata05}
{Ibata}, R.; {Chapman}, S.; {Ferguson}, A.M.N.; {Lewis}, G.; {Irwin}, M.;
  {Tanvir}, N.
\newblock {On the Accretion Origin of a Vast Extended Stellar Disk around the
  Andromeda Galaxy}.
\newblock {\em ApJ} {\bf 2005}, {\em 634},~287--313,
  \href{http://xxx.lanl.gov/abs/astro-ph/0504164}{{\normalfont
  [astro-ph/0504164]}}.

\bibitem[{Toth} and {Ostriker}(1992)]{toth92}
{Toth}, G.; {Ostriker}, J.P.
\newblock {Galactic disks, infall, and the global value of Omega}.
\newblock {\em ApJ} {\bf 1992}, {\em 389},~5--26.

\bibitem[{Quinn} \em{et~al.}(1993){Quinn}, {Hernquist}, and
  {Fullagar}]{quinn93}
{Quinn}, P.J.; {Hernquist}, L.; {Fullagar}, D.P.
\newblock {Heating of galactic disks by mergers}.
\newblock {\em ApJ} {\bf 1993}, {\em 403},~74--93.

\bibitem[{Walker} \em{et~al.}(1996){Walker}, {Mihos}, and
  {Hernquist}]{walker96}
{Walker}, I.R.; {Mihos}, J.C.; {Hernquist}, L.
\newblock {Quantifying the Fragility of Galactic Disks in Minor Mergers}.
\newblock {\em ApJ} {\bf 1996}, {\em 460},~121,
  \href{http://xxx.lanl.gov/abs/astro-ph/9510052}{{\normalfont
  [astro-ph/9510052]}}.

\bibitem[{Font} \em{et~al.}(2001){Font}, {Navarro}, {Stadel}, and
  {Quinn}]{font01}
{Font}, A.S.; {Navarro}, J.F.; {Stadel}, J.; {Quinn}, T.
\newblock {Halo Substructure and Disk Heating in a {$\Lambda$} Cold Dark Matter
  Universe}.
\newblock {\em ApJ Lett} {\bf 2001}, {\em 563},~L1--L4,
  \href{http://xxx.lanl.gov/abs/astro-ph/0106268}{{\normalfont
  [astro-ph/0106268]}}.

\bibitem[{Ardi} \em{et~al.}(2003){Ardi}, {Tsuchiya}, and {Burkert}]{ardi03}
{Ardi}, E.; {Tsuchiya}, T.; {Burkert}, A.
\newblock {Constraints of the Clumpiness of Dark Matter Halos through Heating
  of the Disk Galaxies}.
\newblock {\em ApJ} {\bf 2003}, {\em 596},~204--215,
  \href{http://xxx.lanl.gov/abs/astro-ph/0206026}{{\normalfont
  [astro-ph/0206026]}}.

\bibitem[{Benson} \em{et~al.}(2004){Benson}, {Lacey}, {Frenk}, {Baugh}, and
  {Cole}]{benson04}
{Benson}, A.J.; {Lacey}, C.G.; {Frenk}, C.S.; {Baugh}, C.M.; {Cole}, S.
\newblock {Heating of galactic discs by infalling satellites}.
\newblock {\em MNRAS} {\bf 2004}, {\em 351},~1215--1236,
  \href{http://xxx.lanl.gov/abs/astro-ph/0307298}{{\normalfont
  [astro-ph/0307298]}}.

\bibitem[{Stewart} \em{et~al.}(2008){Stewart}, {Bullock}, {Wechsler}, {Maller},
  and {Zentner}]{stewart08}
{Stewart}, K.R.; {Bullock}, J.S.; {Wechsler}, R.H.; {Maller}, A.H.; {Zentner},
  A.R.
\newblock {Merger Histories of Galaxy Halos and Implications for Disk
  Survival}.
\newblock {\em ApJ} {\bf 2008}, {\em 683},~597--610,
  \href{http://xxx.lanl.gov/abs/0711.5027}{{\normalfont [0711.5027]}}.

\bibitem[{Hopkins} \em{et~al.}(2008){Hopkins}, {Hernquist}, {Cox}, {Younger},
  and {Besla}]{hopkins08}
{Hopkins}, P.F.; {Hernquist}, L.; {Cox}, T.J.; {Younger}, J.D.; {Besla}, G.
\newblock {The Radical Consequences of Realistic Satellite Orbits for the
  Heating and Implied Merger Histories of Galactic Disks}.
\newblock {\em ApJ} {\bf 2008}, {\em 688},~757--769,
  \href{http://xxx.lanl.gov/abs/0806.2861}{{\normalfont [0806.2861]}}.

\bibitem[{Villalobos} and {Helmi}(2008)]{villalobos08}
{Villalobos}, {\'A}.; {Helmi}, A.
\newblock {Simulations of minor mergers - I. General properties of thick
  discs}.
\newblock {\em MNRAS} {\bf 2008}, {\em 391},~1806--1827.

\bibitem[{Purcell} \em{et~al.}(2009){Purcell}, {Kazantzidis}, and
  {Bullock}]{purcell09}
{Purcell}, C.W.; {Kazantzidis}, S.; {Bullock}, J.S.
\newblock {The Destruction of Thin Stellar Disks Via Cosmologically Common
  Satellite Accretion Events}.
\newblock {\em ApJ Lett} {\bf 2009}, {\em 694},~L98--L102,
  \href{http://xxx.lanl.gov/abs/0810.2785}{{\normalfont [0810.2785]}}.

\bibitem[{Kazantzidis} \em{et~al.}(2009){Kazantzidis}, {Zentner}, {Kravtsov},
  {Bullock}, and {Debattista}]{kazantzidis09}
{Kazantzidis}, S.; {Zentner}, A.R.; {Kravtsov}, A.V.; {Bullock}, J.S.;
  {Debattista}, V.P.
\newblock {Cold Dark Matter Substructure and Galactic Disks. II. Dynamical
  Effects of Hierarchical Satellite Accretion}.
\newblock {\em ApJ} {\bf 2009}, {\em 700},~1896--1920,
  \href{http://xxx.lanl.gov/abs/0902.1983}{{\normalfont
  [arXiv:astro-ph.GA/0902.1983]}}.

\bibitem[{Sachdeva} and {Saha}(2016)]{sachdeva16}
{Sachdeva}, S.; {Saha}, K.
\newblock {Survival of Pure Disk Galaxies over the Last 8 Billion Years}.
\newblock {\em ApJ Lett} {\bf 2016}, {\em 820},~L4,
  \href{http://xxx.lanl.gov/abs/1602.08942}{{\normalfont [1602.08942]}}.

\bibitem[{Moetazedian} and {Just}(2016)]{moetazedian16}
{Moetazedian}, R.; {Just}, A.
\newblock {Impact of cosmological satellites on the vertical heating of the
  Milky Way disc}.
\newblock {\em MNRAS} {\bf 2016}, {\em 459},~2905--2924,
  \href{http://xxx.lanl.gov/abs/1508.03580}{{\normalfont [1508.03580]}}.

\bibitem[{Abadi} \em{et~al.}(2006){Abadi}, {Navarro}, and {Steinmetz}]{abadi06}
{Abadi}, M.G.; {Navarro}, J.F.; {Steinmetz}, M.
\newblock {Stars beyond galaxies: the origin of extended luminous haloes around
  galaxies}.
\newblock {\em MNRAS} {\bf 2006}, {\em 365},~747--758,
  \href{http://xxx.lanl.gov/abs/astro-ph/0506659}{{\normalfont
  [astro-ph/0506659]}}.

\bibitem[{Zolotov} \em{et~al.}(2009){Zolotov}, {Willman}, {Brooks},
  {Governato}, {Brook}, {Hogg}, {Quinn}, and {Stinson}]{zolotov09}
{Zolotov}, A.; {Willman}, B.; {Brooks}, A.M.; {Governato}, F.; {Brook}, C.B.;
  {Hogg}, D.W.; {Quinn}, T.; {Stinson}, G.
\newblock {The Dual Origin of Stellar Halos}.
\newblock {\em ApJ} {\bf 2009}, {\em 702},~1058--1067,
  \href{http://xxx.lanl.gov/abs/0904.3333}{{\normalfont
  [arXiv:astro-ph.GA/0904.3333]}}.

\bibitem[{Zolotov} \em{et~al.}(2010){Zolotov}, {Willman}, {Brooks},
  {Governato}, {Hogg}, {Shen}, and {Wadsley}]{zolotov10}
{Zolotov}, A.; {Willman}, B.; {Brooks}, A.M.; {Governato}, F.; {Hogg}, D.W.;
  {Shen}, S.; {Wadsley}, J.
\newblock {The Dual Origin of Stellar Halos. II. Chemical Abundances as Tracers
  of Formation History}.
\newblock {\em ApJ} {\bf 2010}, {\em 721},~738--743,
  \href{http://xxx.lanl.gov/abs/1004.3789}{{\normalfont [1004.3789]}}.

\bibitem[{Font} \em{et~al.}(2011){Font}, {McCarthy}, {Crain}, {Theuns},
  {Schaye}, {Wiersma}, and {Dalla Vecchia}]{font11}
{Font}, A.S.; {McCarthy}, I.G.; {Crain}, R.A.; {Theuns}, T.; {Schaye}, J.;
  {Wiersma}, R.P.C.; {Dalla Vecchia}, C.
\newblock {Cosmological simulations of the formation of the stellar haloes
  around disc galaxies}.
\newblock {\em MNRAS} {\bf 2011}, {\em 416},~2802--2820,
  \href{http://xxx.lanl.gov/abs/1102.2526}{{\normalfont [1102.2526]}}.

\bibitem[{McCarthy} \em{et~al.}(2012){McCarthy}, {Font}, {Crain}, {Deason},
  {Schaye}, and {Theuns}]{mccarthy12}
{McCarthy}, I.G.; {Font}, A.S.; {Crain}, R.A.; {Deason}, A.J.; {Schaye}, J.;
  {Theuns}, T.
\newblock {Global structure and kinematics of stellar haloes in cosmological
  hydrodynamic simulations}.
\newblock {\em MNRAS} {\bf 2012}, {\em 420},~2245--2262,
  \href{http://xxx.lanl.gov/abs/1111.1747}{{\normalfont
  [arXiv:astro-ph.GA/1111.1747]}}.

\bibitem[{Tissera} \em{et~al.}(2013){Tissera}, {Scannapieco}, {Beers}, and
  {Carollo}]{tissera13}
{Tissera}, P.B.; {Scannapieco}, C.; {Beers}, T.C.; {Carollo}, D.
\newblock {Stellar haloes of simulated Milky-Way-like galaxies: chemical and
  kinematic properties}.
\newblock {\em MNRAS} {\bf 2013}, {\em 432},~3391--3400,
  \href{http://xxx.lanl.gov/abs/1301.1301}{{\normalfont [1301.1301]}}.

\bibitem[{Tissera} \em{et~al.}(2014){Tissera}, {Beers}, {Carollo}, and
  {Scannapieco}]{tissera14}
{Tissera}, P.B.; {Beers}, T.C.; {Carollo}, D.; {Scannapieco}, C.
\newblock {Stellar haloes in Milky Way mass galaxies: from the inner to the
  outer haloes}.
\newblock {\em MNRAS} {\bf 2014}, {\em 439},~3128--3138,
  \href{http://xxx.lanl.gov/abs/1309.3609}{{\normalfont
  [arXiv:astro-ph.CO/1309.3609]}}.

\bibitem[{Pillepich} \em{et~al.}(2015){Pillepich}, {Madau}, and
  {Mayer}]{pillepich15}
{Pillepich}, A.; {Madau}, P.; {Mayer}, L.
\newblock {Building Late-type Spiral Galaxies by In-situ and Ex-situ Star
  Formation}.
\newblock {\em ApJ} {\bf 2015}, {\em 799},~184,
  \href{http://xxx.lanl.gov/abs/1407.7855}{{\normalfont [1407.7855]}}.

\bibitem[{Cooper} \em{et~al.}(2015){Cooper}, {Parry}, {Lowing}, {Cole}, and
  {Frenk}]{cooper15}
{Cooper}, A.P.; {Parry}, O.H.; {Lowing}, B.; {Cole}, S.; {Frenk}, C.
\newblock {Formation of in situ stellar haloes in Milky Way-mass galaxies}.
\newblock {\em MNRAS} {\bf 2015}, {\em 454},~3185--3199,
  \href{http://xxx.lanl.gov/abs/1501.04630}{{\normalfont [1501.04630]}}.

\bibitem[{Carollo} \em{et~al.}(2007){Carollo}, {Beers}, {Lee}, {Chiba},
  {Norris}, {Wilhelm}, {Sivarani}, {Marsteller}, {Munn}, {Bailer-Jones},
  {Fiorentin}, and {York}]{carollo07}
{Carollo}, D.; {Beers}, T.C.; {Lee}, Y.S.; {Chiba}, M.; {Norris}, J.E.;
  {Wilhelm}, R.; {Sivarani}, T.; {Marsteller}, B.; {Munn}, J.A.;
  {Bailer-Jones}, C.A.L.; {Fiorentin}, P.R.; {York}, D.G.
\newblock {Two stellar components in the halo of the Milky Way}.
\newblock {\em Nature} {\bf 2007}, {\em 450},~1020--1025,
  \href{http://xxx.lanl.gov/abs/0706.3005}{{\normalfont [0706.3005]}}.

\bibitem[{Deason} \em{et~al.}(2013){Deason}, {Belokurov}, {Evans}, and
  {Johnston}]{deason13}
{Deason}, A.J.; {Belokurov}, V.; {Evans}, N.W.; {Johnston}, K.V.
\newblock {Broken and Unbroken: The Milky Way and M31 Stellar Halos}.
\newblock {\em ApJ} {\bf 2013}, {\em 763},~113,
  \href{http://xxx.lanl.gov/abs/1210.4929}{{\normalfont [1210.4929]}}.

\bibitem[{Dorman} \em{et~al.}(2013){Dorman}, {Widrow}, {Guhathakurta}, {Seth},
  {Foreman-Mackey}, {Bell}, {Dalcanton}, {Gilbert}, {Skillman}, and
  {Williams}]{dorman13}
{Dorman}, C.E.; {Widrow}, L.M.; {Guhathakurta}, P.; {Seth}, A.C.;
  {Foreman-Mackey}, D.; {Bell}, E.F.; {Dalcanton}, J.J.; {Gilbert}, K.M.;
  {Skillman}, E.D.; {Williams}, B.F.
\newblock {A New Approach to Detailed Structural Decomposition from the SPLASH
  and PHAT Surveys: Kicked-up Disk Stars in the Andromeda Galaxy?}
\newblock {\em ApJ} {\bf 2013}, {\em 779},~103,
  \href{http://xxx.lanl.gov/abs/1310.4179}{{\normalfont [1310.4179]}}.

\bibitem[{Hawkins} \em{et~al.}(2015){Hawkins}, {Kordopatis}, {Gilmore},
  {Masseron}, {Wyse}, {Ruchti}, {Bienaym{\'e}}, {Bland-Hawthorn}, {Boeche},
  {Freeman}, {Gibson}, {Grebel}, {Helmi}, {Kunder}, {Munari}, {Navarro},
  {Parker}, {Reid}, {Scholz}, {Seabroke}, {Siebert}, {Steinmetz}, {Watson}, and
  {Zwitter}]{hawkins15}
{Hawkins}, K.; {Kordopatis}, G.; {Gilmore}, G.; {Masseron}, T.; {Wyse}, R.F.G.;
  {Ruchti}, G.; {Bienaym{\'e}}, O.; {Bland-Hawthorn}, J.; {Boeche}, C.;
  {Freeman}, K.; {Gibson}, B.K.; {Grebel}, E.K.; {Helmi}, A.; {Kunder}, A.;
  {Munari}, U.; {Navarro}, J.F.; {Parker}, Q.A.; {Reid}, W.A.; {Scholz}, R.D.;
  {Seabroke}, G.; {Siebert}, A.; {Steinmetz}, M.; {Watson}, F.; {Zwitter}, T.
\newblock {Characterizing the high-velocity stars of RAVE: the discovery of a
  metal-rich halo star born in the Galactic disc}.
\newblock {\em MNRAS} {\bf 2015}, {\em 447},~2046--2058,
  \href{http://xxx.lanl.gov/abs/1412.1484}{{\normalfont [1412.1484]}}.

\bibitem[{Bonaca} \em{et~al.}(2017){Bonaca}, {Conroy}, {Wetzel}, {Hopkins}, and
  {Keres}]{bonaca17}
{Bonaca}, A.; {Conroy}, C.; {Wetzel}, A.; {Hopkins}, P.F.; {Keres}, D.
\newblock {Gaia reveals a metal-rich in-situ component of the local stellar
  halo}.
\newblock {\em ArXiv e-prints} {\bf 2017},
  \href{http://xxx.lanl.gov/abs/1704.05463}{{\normalfont [1704.05463]}}.

\bibitem[{Chakrabarti} and {Blitz}(2009)]{chakrabarti09}
{Chakrabarti}, S.; {Blitz}, L.
\newblock {Tidal imprints of a dark subhalo on the outskirts of the Milky Way}.
\newblock {\em MNRAS} {\bf 2009}, {\em 399},~L118--L122,
  \href{http://xxx.lanl.gov/abs/0812.0821}{{\normalfont [0812.0821]}}.

\bibitem[{Chakrabarti} \em{et~al.}(2011){Chakrabarti}, {Bigiel}, {Chang}, and
  {Blitz}]{chakrabarti11b}
{Chakrabarti}, S.; {Bigiel}, F.; {Chang}, P.; {Blitz}, L.
\newblock {Finding Dwarf Galaxies from Their Tidal Imprints}.
\newblock {\em ApJ} {\bf 2011}, {\em 743},~35,
  \href{http://xxx.lanl.gov/abs/1101.0815}{{\normalfont [1101.0815]}}.

\bibitem[{Chang} and {Chakrabarti}(2011)]{chang11}
{Chang}, P.; {Chakrabarti}, S.
\newblock {Dark subhaloes and disturbances in extended H I discs}.
\newblock {\em MNRAS} {\bf 2011}, {\em 416},~618--628,
  \href{http://xxx.lanl.gov/abs/1102.3436}{{\normalfont [1102.3436]}}.

\bibitem[Duc et al.(2015)]{duc15} Duc, P.-A., and 26 colleagues 2015.\ The ATLAS$^{3D}$ project - XXIX. The new look of early-type galaxies and surrounding fields disclosed by extremely deep optical images.\ Monthly Notices of the Royal Astronomical Society 446, 120-143.

\bibitem[Mart{\'{\i}}nez-Delgado et al.(2010)]{delgado10} Mart{\'{\i}}nez-Delgado, D., and 14 colleagues 2010.\ Stellar Tidal Streams in Spiral Galaxies of the Local Volume: A Pilot Survey with Modest Aperture Telescopes.\ The Astronomical Journal 140, 962-967.

\bibitem[van Dokkum et al.(2014)]{vandokkum14} van Dokkum, P.~G., Abraham, R., Merritt, A.\ 2014.\ First Results from the Dragonfly Telephoto Array: The Apparent Lack of a Stellar Halo in the Massive Spiral Galaxy M101.\ The Astrophysical Journal 782, L24.

\bibitem[Spergel et al.(2013)]{spergel13} Spergel, D., and 45 colleagues 2013.\ WFIRST-2.4: What Every Astronomer Should Know.\ ArXiv e-prints arXiv:1305.5425.

\bibitem[Ivezic et al.(2008)]{ivezic08} Ivezic, Z., and 26 colleagues 2008.\ Large Synoptic Survey Telescope: From Science Drivers To Reference Design.\ Serbian Astronomical Journal 176, 1-13.

\bibitem[Monachesi et al.(2013)]{monachesi13} Monachesi, A., Bell, E.~F., Radburn-Smith, D.~J., Vlaji{\'c}, M., de Jong, R.~S., Bailin, J., Dalcanton, J.~J., Holwerda, B.~W., Streich, D.\ 2013.\ Testing Galaxy Formation Models with the GHOSTS Survey: The Color Profile of M81's Stellar Halo.\ The Astrophysical Journal 766, 106.

\bibitem[Crnojevi{\'c} et al.(2016)]{crnojevi16} Crnojevi{\'c}, D., Sand, D.~J., Spekkens, K., Caldwell, N., Guhathakurta, P., McLeod, B., Seth, A., Simon, J.~D., Strader, J., Toloba, E.\ 2016.\ The Extended Halo of Centaurus A: Uncovering Satellites, Streams, and Substructures.\ The Astrophysical Journal 823, 19.

\bibitem[D'Onghia et al.(2016)]{donghia16} D'Onghia, E., Madau, P., Vera-Ciro, C., Quillen, A., Hernquist, L.\ 2016.\ Excitation of Coupled Stellar Motions in the Galactic Disk by Orbiting Satellites.\ The Astrophysical Journal 823, 4.

\bibitem[Crane et al.(2003)]{crane03} Crane, J.~D., Majewski, S.~R., Rocha-Pinto, H.~J., Frinchaboy, P.~M., Skrutskie, M.~F., Law, D.~R.\ 2003.\ Exploring Halo Substructure with Giant Stars: Spectroscopy of Stars in the Galactic Anticenter Stellar Structure.\ The Astrophysical Journal 594, L119-L122.

\bibitem[Chou et al.(2011)]{chou11} Chou, M.-Y., Majewski, S.~R., Cunha, K., Smith, V.~V., Patterson, R.~J., Mart{\'{\i}}nez-Delgado, D.\ 2011.\ First Chemical Analysis of Stars in the Triangulum--Andromeda Star Cloud.\ The Astrophysical Journal 731, L30.

\bibitem[Sheffield et al.(2012)]{sheffield12} Sheffield, A.~A., and 12 colleagues 2012.\ Identifying Contributions to the Stellar Halo from Accreted, Kicked-out, and In Situ Populations.\ The Astrophysical Journal 761, 161.

\bibitem[Velazquez and White(1999)]{velazquez99} Velazquez, H., White, S.~D.~M.\ 1999.\ Sinking satellites and the heating of galaxy discs.\ Monthly Notices of the Royal Astronomical Society 304, 254-270.

%\bibitem[Kazantzidis et al.(2008)]{kazantzidis08} Kazantzidis, S., Bullock, J.~S., Zentner, A.~R., Kravtsov, A.~V., Moustakas, L.~A.\ 2008.\ Cold Dark Matter Substructure and Galactic Disks. I. Morphological Signatures of Hierarchical Satellite Accretion.\ The Astrophysical Journal 688, 254-276.

\bibitem[Deng et al.(2012)]{deng12} Deng, L.-C., and 25 colleagues 2012.\ LAMOST Experiment for Galactic Understanding and Exploration (LEGUE) {---} The survey's science plan.\ Research in Astronomy and Astrophysics 12, 735-754.

\bibitem[Laporte et al.(2016)]{laporte16} Laporte, C.~F.~P., G{\'o}mez, F.~A., Besla, G., Johnston, K.~V., Garavito-Camargo, N.\ 2016.\ Response of the Milky Way's disc to the Large Magellanic Cloud in a first infall scenario.\ ArXiv e-prints arXiv:1608.04743.

\bibitem[Li et al.(2017)]{li17} Li, T.~S., and 10 colleagues 2017.\ Exploring Halo Substructure with Giant Stars. XV. Discovery of a Connection between the Monoceros Ring and the Triangulum-Andromeda Overdensity?.\ ArXiv e-prints arXiv:1703.05384.

\end{thebibliography}